Stack Safety as a Security Property

### 2021-04-02 tex/main.tex
Work on examples
Extended examples a bit more
More work on examples and intro

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\input{macros}
\input{examples}

@@

\newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}

@@

But what {\em is} this stack safety of which they speak? Precise formal
definitions are mostly lacking: rather, the notion is commonly defined in a
negative way, by enumerating the bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation} and {\em well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety. In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.'' While our criteria are
designed with PIPE micro-policies in mind, they are agnostic about many implementation
details and should be applicable to other systems with minor adjustment.
%
Our key contribution is an abstract presentation of local state encapsulation
using the technical framework of language-based security, together with a simple
characterization of well-bracketed control flow.

\begin{figure}[b]
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 42;
  f();
  print(x);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 1;
  f();
  if(x)
    g();
  else
    h();
}
\end{verbatim}
    \end{minipage}
  \caption{Simple Examples}
  \label{fig:examples1}
\end{figure}

\begin{figure}[b]
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main(int arg) {
  if(arg)
    g();
  else
    h();
  f();
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 1;
  f();
  x = -x;
  f();
  print(x);
}
\end{verbatim}
  \end{minipage}

  \caption{More Examples}
  \label{fig:examples2}
\end{figure}

@@

When does the programmer rely on stack safety? The most natural example is the first
in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
overwrite {\tt main}'s stack data. This expectation is for {\em integrity}: that a
caller's data is protected from modification by its callee.
This is not only useful for the programmer: consider the second
example, where the call to {\tt h} should be unreachable and therefore subject to dead code
elimination, if only the compiler can soundly reason that {\tt x} does not change.

Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
from leakage by its callee. Consider again the first example in figure \ref{fig:examples1},
in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
programmer had initialized {\tt x} after calling {\tt f} rather than before; such a change
should not alter the behavior of {\tt main}, but if {\tt f} can directly read {\tt x}, it
may. But there are more subtle examples, such as the first in figure \ref{fig:examples2},
where {\tt main} may call {\tt g} or {\tt h} are called prior to calling {\tt f}, depending on
its argument. Here, if {\tt g} or {\tt h} leaves behind identifying information in unused stack
space, {\tt f} could still deduce information about the argument and behave differently, violating
the abstraction of its call.

Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
callers from their callees. Consider the last example in figure \ref{fig:examples2},
a variant of the famous ``awkward example,'' in which {\tt f} is called twice and might store
data outside the stack between calls. It should be the case that {\tt main} prints -1, but if
{\tt f}'s first activation stores its return address somewhere accessible, the second
activation can return to before {\tt x} is negated, causing {\tt x} to be set to
1 again before a third activation that would return normally. Then {\tt main} would print 1.
This is rather unexpected behavior.

Preventing this form of control-flow attack is the intuition behind
\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
the call graph itself. A callee should pass control back to the caller only by
transferring control to the instruction following the call; otherwise it could
subvert LSE by appearing to return -- thus restoring access to the caller's data --
while causing the execution of arbitrary instructions on that data.

@@

First we introduce this model for a single simple stack with no sharing between callers and
callee. Then we expand it to allow components to be passed, granting a callee one time
access to the caller's data, or shared, allowing components to be outright
exempted from the stack safety properties. Shared components could model objects whose
addresses are passed and whose security must be modeled by a separate property, such as
memory safety enforced by capabilities. Finally, we extend the model to a coroutine
system featuring finitely-bounded stacks.

After formally defining stack safety, we show that the definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
This policy testably enforces our {\em eager} properties, in which each
callee, at the point of its return, can be seen to respect the integrity
and confidentiality of its caller. Additionally we have an
{\em ultra-eager} integrity property which guarantees that integrity
is respected at each step. The advantage of this property is that it ``fails
faster'' and is therefore more suitable for efficient testing. We prove (in
Coq) that the ultra-eager property implies our original definition of integrity.

The downside of the Depth Isolation policy is that it is rather
expensive to implement. In the same work,
\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment when a callee returns, this policy signals violations when the
caller actually accesses data that the callee has written into its stack
frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property in the same way...}\fi
The lazy policy does not enforce our original integrity property, and to our
knowledge there is no formal statement of the properties that it ought to enforce.
We propose a {\em lazy} integrity property, which capture the desired behavior of a
policy that detects integrity violations between when they occur and when they make
a visible impact. Intuitively this can be phrased, ``No visible reads after dangerous
writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
``Lazy Tagging and Clearing'' policy does not quite enforce this property, as shown in
testing, but formalizing the ideal property allows us to see how it falls short.
We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
at some performance cost.

@@

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
  A novel {\em stack-safety property} abstractly capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}).
  % These properties rule out common examples of stack memory abuse.
\item
  A property-based randomized testing framework that demonstrates that this
  property can be enforced by the Depth Isolation policy of
  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware tagging~(\cref{sec:conservative}).
\item
  An {\em ultra-eager} variant of the stack-integrity property that is also
  enforced by the Depth Isolation policy but is more efficient to test, plus a proof
  (in Coq) that this variant is stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced if the
  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
  with unique tags on each activation, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
\item
  Standardized trace-based formulations of eager confidentiality and of ultra-eager,
  eager, and lazy integrity, parameterizeable by arbitrary control-flow models and assignments
  of machine components to {\em LSE domains}.
\item
  An LSE-domain model tracking the passing of function arguments on the stack and the sharing
  of arbitrary stack data.
\item
  An extension of stack safety to a simple coroutine model with corresponding LSE
  properties and a well-structured control flow property that subsumes WBCF for each
  stack as well as guaranteeing legitimate coroutine crossings.
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

\paragraph{Well-formedness and Bad Annotations}

@@

\paragraph{System Assumptions}

@@

\paragraph{Policies}

@@

\section{Examples and Intuition}
\label{sec:running-example}

\input{example1}

\input{example2}

@@

    \begin{figure}
      \integritylazyexample
      \caption{Rolling back from Figure \ref{fig:intex}.}
        %\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
      \label{fig:intlex}
    \end{figure}

@@

      \begin{figure}
        \confidentialitylazyexample
        \caption{Continuing from \cref{fig:confex}.}
        \label{fig:conflex}
      \end{figure}

@@

\section{Extension to Software-based Enforcement}

>>>>

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\input{macros}
\input{examples}
\input{example1}

@@

\newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}
\newcommand{\memoryaddrs}[1]
  {
    \MemoryLabel{#1}{2em}{\SP}
  }

@@

But what {\em is} this stack safety of which they speak? Precise formal
definitions are mostly lacking: rather, the notion is commonly defined in a
negative way, by enumerating the bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation} and {\em well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety. In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.'' We begin with a strong
local state encapsulation property that closely models the behavior of eager
micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
is too strong to characterize stack safety; reasonable enforcement mechanisms
exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
to support this technique and argue that the new property, defined relative to an
arbitrary notion of observation, is weak enough to characterize stack safety.

While our criteria are designed with PIPE micro-policies in mind, they are
agnostic about many implementation details and should be applicable to other
systems with minor adjustment.
%
Our key contribution is an operational model of local state encapsulation
using the technical framework of language-based security, with eager and lazy variants,
together with a simple characterization of well-bracketed control flow.

%\begin{figure}[b]
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main() {
%  int x = 42;
%  f();
%  print(x);
%}
%\end{verbatim}
%    \end{minipage}
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main() {
%  int x = 1;
%  f();
%  if(x)
%    g();
%  else
%    h();
%}
%\end{verbatim}
%    \end{minipage}
%  \caption{Simple Examples}
%  \label{fig:examples1}
%\end{figure}

%\begin{figure}[b]
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main(int arg) {
%  if(arg)
%    g();
%  else
%    h();
%  f();
%}
%\end{verbatim}
%    \end{minipage}
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main() {
%  int x = 1;
%  f();
%  x = -x;
%  f();
%  print(x);
%}
%\end{verbatim}
%  \end{minipage}
%
%  \caption{More Examples}
%  \label{fig:examples2}
%\end{figure}

@@

How does the programmer rely on stack safety? The stack makes implicit promises about
the behavior of function calls. First, that when a call is finished, the caller can continue
from where it left off with its local state unchanged. And second, that the call is
self-contained, its behavior determined only by a predictable interface -- at minimum
parameters and global variables, but never local state.

%The most natural example is the first
%in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
%calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
%{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
%overwrite {\tt main}'s stack data.

Implicit in both expectations are a predictable control flow, termed {\em well-bracketed
control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls jump to a consistent entry
point and return to the call site. Then the data portion of the first expectation is for
{\em integrity}: that a caller's local data is protected from modification by its callee.
This is vital for local reasoning about the behavior of the caller. The security implications
are significant: a caller that branches on a local variable could be manipulated by its
callee into unexpected behavior without integrity.

Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
from leakage by its callee. As a reasoning principle, confidentiality tells us that to
predict the behavior of a call we need only consider a subset of the machine state.
Rearranging local data will have no impact. And in security terms, we can guarantee
the secrecy of local data that we haven't passed even when calling a badly behaved function.
Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
{\em local state encapsulation} (LSE) property.

It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
accessing the caller's state, but the execution would remain under the callee's control.

@@

First we introduce this model for a single simple stack with no sharing between callers and
callee. Then we expand it to allow components to be passed, granting a callee one time
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (The combination of such policies is
a strength of tag-based approaches.)

Finally, we extend the model to a coroutine system featuring finitely-bounded stacks. Though
simple, this extension is a step toward concurrency and demonstrates that the model is flexible
enough to handle more sophisticated styles of control-flow.

After formally defining stack safety, we show that the definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
This policy testably enforces our {\em eager} properties, in which integrity is respected
at every step of a program, and each call respects integrity at each of its internal steps.
These properties are very strong, and our policies enforce them; they also have the advantage
that they ``fail faster'' and are therefore suitable for efficient testing.

The downside of the Depth Isolation policy is that it is rather expensive to implement.
In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment when a callee returns, this policy signals violations when the
caller actually accesses data that the callee has written into its stack
frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property in the same way...}\fi
The lazy policy does not enforce our eager integrity property, and to our
knowledge there is no formal statement of the properties that it ought to enforce.
We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
policy that detects integrity violations between when they occur and when they make
a visible impact. Intuitively this can be phrased, ``No visible reads after dangerous
writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
testing, but formalizing the ideal property allows us to see how it falls short.
We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
at some performance cost. We prove (in Coq) that the eager property implies the lazy one.
We argue that any system purporting to enforce stack safety ought to obey lazy stack safety.

@@

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
  A novel {\em stack-safety property} capturing the intuitive concepts of strictly enforced
  local state encapsulation and well-bracketed control flow~(\cref{sec:lse-and-wbcf}).
  % These properties rule out common examples of stack memory abuse.
\item
  A property-based randomized testing framework that demonstrates that this
  property can be enforced by the Depth Isolation policy of
  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware
  tagging~(\cref{sec:conservative}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced if the
  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
  with unique tags on each activation, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
\item
  Standardized trace-based formulations of both eager and lazy confidentiality and integrity,
  parameterizeable by arbitrary control-flow models and assignments of machine components
  to {\em LSE domains}.
\item
  An LSE-domain model enhanced with passing of function arguments on the stack, with which our
  standardized trace properties extend stack safety with local sharing.
\item
  An LSE-domain model extending the stack to a simple coroutine model, with corresponding LSE
  properties and a {\em well-structured control flow} property that subsumes WBCF for each
  stack as well as guaranteeing legitimate coroutine crossings.
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

\paragraph*{Well-formedness and Bad Annotations}

@@

\paragraph*{System Assumptions}

@@

\paragraph*{Policies}

@@

\section{Stack Safety, Intuitively}
\label{sec:running-example}
\colorlet{lgray}{gray!40}

\newcommand{\mainsealc}{cyan}
\newcommand{\fsealc}{green}
\newcommand{\unsealc}{lgray}
\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
\newcommand{\fulloutc}{white}

\begin{figure}

\begin{subfigure}{.4\textwidth}
\begin{verbatim}
int main() {
0:     int x = 42, y = 0;
1[C]:  y = f(10);
2:     print (x + y);
}

int f(int z) {
10:     print z;
11[R]:  return (z+z);
}
\end{verbatim}

\caption{A Simple Call}
\label{fig:simple-program}
\end{subfigure}
\begin{subfigure}{.4\textwidth}  
\begin{center}
\begin{tabular}{l l}
{\tt 0} &
\memoryaddrs{4em}
\memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
~$\cdots$
\\
{\tt 1} &
\memoryaddrs{12em}
\memory{3}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{42}
\MemoryLabel{-10em}{0.75em}{0}
\vspace{.5em}
\\
{\tt 10} &
\memoryaddrs{17em}
\memory{2}{\mainsealc}[{\makebox[0pt]{Sealed(0)}}]%
\memory{2}{\unsealc}
~$\cdots$
\MemoryLabel{-10em}{0.75em}{2}
\\
{\tt 2} &
\memoryaddrs{12em}
\memory{3}{\unsealc}
~$\cdots$
\MemoryLabel{-14em}{0.75em}{42}
\MemoryLabel{-10em}{0.75em}{0}
\MemoryLabel{-7em}{0.75em}{2}
\\
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}

\caption{Trace of Stack Domains}

\label{fig:simple-trace}
\end{subfigure}
\end{figure}

This section will elucidate the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will insert our model between the enforcement and
the examples, arguing that:

\begin{itemize}
\item Examples that appear to violate stack safety, violate at least one of our properties
\item We do not find desireable behaviors that appear stack-safe yet violate our properties
\item Our properties are testably implemented by stack safety micro-policies from the literature,
  and the micro-policies are not significantly more conservative (i.e., policy failstops tend to
  correspond to imminent property violations)
\end{itemize}

\Cref{fig:simple-program} shows a simple example, written in C-like pseudo-code.
It consists of a {\tt main} function that performs a call to a function {\tt f},
which prints its argument, performs arithmetic, and returns a result.
\Cref{fig:simple-trace} shows the stack as it is divided into domains throughout
the program execution. We use high-level code for compactness, but reason about
a standard compilation of each function. Line numbers correspond to the code
addresses just before the code that implements the line, and annotations
{\tt C} and {\tt R} indicate that an instruction is annotated as a call or
return, respectively.

We also assume that the machine is word- (not byte-) addressed and that the
machine communicates with the outside world through a single memory-mapped
output port whose address is elided (so a {\tt print} statement will be
compiled to a store to this location), and that the stack grows upwards.

In the initial memory layout the entire stack is currently {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals (as designated by the stack pointer) are
{\em sealed} and labeled with the depth of their owner, in this case 0.
Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. Later we will
loosen this condition in the interest of extensionality -- the caller only
cares about changes it can see after the return, not changes during execution.

Suppose we insert an instruction after line 10, that stores 5 to the stack pointer - 1.
Then we would be changing {\tt y}, which is sealed, in violation of stepwise integrity.

\begin{figure}

\begin{subfigure}{.33\textwidth}
\begin{verbatim}
int main() {
0:     int x = 42, y = 0;
1[C]:  y = f();
2:     print (x + y);
}

int f() {
10:     int z = *($SP-1);
11[R]:  return (z+z);
}
\end{verbatim}
\label{fig:conf1-program}
\end{subfigure}
\begin{subfigure}{.59\textwidth}
\centering
\begin{tabular}{l l l}
  {\tt 0} &
  \multicolumn{2}{c}{
    \memoryaddrs{4em}
    \memory{3}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-14em}{0.75em}{a}
    \MemoryLabel{-10em}{0.75em}{b}
    \MemoryLabel{-7em}{0.75em}{c}
    \vspace{.5em}
  } \\
  {\tt 1} &
  \multicolumn{2}{c}{
    \memoryaddrs{12em}
    \memory{3}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-14em}{0.75em}{42}
    \MemoryLabel{-10em}{0.75em}{0}
    \MemoryLabel{-7em}{0.75em}{c}
    \vspace{.5em}
  } \\
  {\tt 10} &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf d}
 &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{a'}
  \MemoryLabel{-14em}{0.75em}{b'}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf d'}
  \\
  {\tt 11} &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf 0}
  &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{a'}
  \MemoryLabel{-14em}{0.75em}{b'}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf b'}
  \\
\end{tabular}

\vspace{\abovedisplayskip}

\label{fig:conf1-trace}
\end{subfigure}
\caption{A Confidentiality Violation}
\end{figure}

Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
of the context of its call (with the exception, of course, of its arguments).
We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
security. Noninterference describes knowledge: it states that for some component(s)
that we do not know, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the components by observing those results, violating
confidentiality.

In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt y}, which was not passed to it. The trace
in figure \ref{fig:conf1-trace} shows how noninterference works. First, let \(a\),
\(b\), and \(c\) be the unknown initial values at their locations. On entry to {\tt f}
we again have 42 and 0 in {\tt x} and {\tt y}, and {\tt z} contains \(c\). We consider a
{\em variant state} that replaces the values of components that should be secret with
arbitrary values \(a'\), \(b'\), and \(c'\). After a step, we've copied 0 and \(c'\),
respectively, into {\tt z}. Since {\tt z} changed value, for confidentiality to hold,
its value should be the same in all variant states. But when \(c' \not = 0\), the
states changes differently, which violates confidentiality. It would also be a violation
to print the secret value.

This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. Any components
whose values update and any printed values will be identical between the primary and variant
traces.

\begin{figure}

\begin{subfigure}{.25\textwidth}
\begin{verbatim}
int main() {
0[C]:  int t = f();
1[R]:  return;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.37\textwidth}
\begin{verbatim}
int f() {
10:     int x, y = 5, z = 6;
11:     print y+z;
12:     $SP-=2;
13[C]:  x = g();
14[R]:  return x;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.32\textwidth}
\begin{verbatim}
int g() {
20:     int z;
21[R]:  return (z+z);
}
\end{verbatim}
\end{subfigure}

\begin{subfigure}{.8\textwidth}  
\begin{center}
\begin{tabular}{l l l}
  {\tt 13} &
  \multicolumn{2}{c}{
    \memoryaddrs{16em}
    \memory{1}{\mainsealc}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-23em}{0.75em}{a}
    \MemoryLabel{-19em}{0.75em}{b}
    \MemoryLabel{-15em}{0.75em}{1}
    \MemoryLabel{-11em}{0.75em}{5}
    \MemoryLabel{-8em}{0.75em}{6}
    \vspace{.5em}
  }
  &
\\
  {\tt 20} &
  \memoryaddrs{25em}
  \memory{1}{\mainsealc}
  \memory{2}{\fsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-23em}{0.75em}{a}
  \MemoryLabel{-19em}{0.75em}{b}
  \MemoryLabel{-15em}{0.75em}{1}
  \MemoryLabel{-11em}{0.75em}{14}
  \MemoryLabel{-8em}{0.75em}{\bf 6}
  \vspace{.5em}
  &
  \memoryaddrs{25em}
  \memory{1}{\mainsealc}
  \memory{2}{\fsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-23em}{0.75em}{a'}
  \MemoryLabel{-19em}{0.75em}{b'}
  \MemoryLabel{-15em}{0.75em}{c'}
  \MemoryLabel{-11em}{0.75em}{14}
  \MemoryLabel{-8em}{0.75em}{\bf d'}
  \vspace{.5em}
\\
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}

\end{subfigure}

\caption{Another Confidentiality Violation}
\label{fig:conf2}
\end{figure}

Two important details of confidentiality are reflected in figure \ref{fig:conf2}.
First of all, it has multiple calls, and while we focus on the stack at the call
from line 13 to line 20, it should be clear that we also need to consider a similar
variant at the call from 0 to 10. Confidentiality is in this sense recursive: each
nested call must obey confidentiality in its own right.

More importantly, careful readers will have noticed that the informal definition
of stack confidentiality refers to a variation of the entire stack, not just the
sealed portion. In fact, we support temporal stack safety, in which values left behind
by previous activations are also inaccessible. In this case, we connect temporal safety
firmly to the caller-callee attacker model -- {\tt f} stored some data in the variables
{\tt y} and {\tt z}, used it, then deallocated them by moving the stack pointer.
After the stack pointer is written over {\tt y}, the value of {\tt z} is still
left behind. Then {\tt g} can read it just by allocating and accessing an uninitialized
variable!

\paragraph*{Control Flow and Awkward Examples}

\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:     int x = 1;
1[C]:  f();
2:     x = -x;
3[C]:  f();
4[R]:  return x;
}

void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12:     } else {
13:       $SP-1 = stash;
14:       stash = 0;
15[R]:    return;
        }
}
\end{verbatim}
  \end{subfigure}
\label{fig:WBCF1}
  \begin{subfigure}[t]{.4\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:     int x = 1;
1[C]:  f();
2:     x = -x;
3[R]:  return x;
}

void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12[C]     main();
12:     } else {
13:       $SP-1 = stash;
14:       stash = 0;
15[R]:    return;
        }
}
\end{verbatim}
\label{fig:WBCF2}
  \end{subfigure}
  \caption{Control Flow Attacks}
  \label{fig:WBCF}
\end{figure}

Example \ref{fig:WBCF1} shows how control flow attacks can compromise integrity in principle
even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
global variable and retrieves it the second time it is called, returning to the first call
point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.

To prevent this, we define Well-bracketed Control Flow. Throughout the above examples, we
have implicitly tracked the depth of the current call in order to seal memory at the correct
depth. We also have a clear sense of where the entry points to functions are -- the top line,
in our examples, and another instruction annotation in practice. Well-bracketed Control Flow
holds if:

\begin{itemize}
\item All calls jump to entry points ({\em entry integrity})
\item The depth counter decreases by exactly one on a return ({\em non-skipping})
\item From a call at depth \(d\), the next state to have depth \(d\) matches the stack pointer,
  and the program counter is at the instruction following the jal ({\em return integrity})
\end{itemize}

Non-skipping comes into play for attacks such as seen in \ref{WBCF2}, where the calls are
nested rather than sequential. \sna{The example isn't quite right; need to look at the
  Towards Complete... paper again.}

\sna{Put sharing example here, or after the formalism?}

@@

%    \begin{figure}
%      \integritylazyexample
%      \caption{Rolling back from Figure \ref{fig:intex}.}
        %\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
%      \label{fig:intlex}
%    \end{figure}

@@

%      \begin{figure}
%        \confidentialitylazyexample
%        \caption{Continuing from \cref{fig:confex}.}
%        \label{fig:conflex}
%      \end{figure}

@@

%\section{Extension to Software-based Enforcement}
